---
title: "Altmetric Articles"
date: 2018-07-24
output: 
  html_notebook: default
  github_document: default
---

```{r setup, include = FALSE}
library(tidyverse)
articles <- read_csv('../files_out/20180411_1510_master.csv')

```

# Article data

The *article* is the master Altmetric entity, and contains all the basic information about title, publication date, the first author etc, as well as summaries, such as the numbers of mentions, the altmetric scores and so on. 

Most of the fun to be had with article data is achieved by joining it to other datasets, but how to do so and what can be achieved are described in the notebooks for those datasets. See the Notebook_Subjects.rmd for an example.


```{r Plot of Altmetric scores over time}
ggplot(data = articles) + 
  geom_point(mapping = aes(
    x = print_publication_date, 
    y = altmetric_score
    ))
```

## Article data dictionary

Articles have the following fields:

### altmetric_id

An article identifier used internally by Altmetric themselves, but included here in case there are reasons to contact Altmetric about anything that looks odd.

### altmetric_score

The score altmetric gives to articles depending upon the amount of attention they have received. This is [described in more detail by Altmetric themselves](https://help.altmetric.com/support/solutions/articles/6000059309-about-altmetric-and-the-altmetric-attention-score).

```{r Arranging articles by Altmetric score}

arrange(select(articles, 
               article_title,
               first_author,
               altmetric_score), 
        desc(altmetric_score))

```

### article title

**Data type: character**

The (often very long - go figure) title of the article.

### doi (Digital Object Identifier)

**Data type: character**

In the current version of the Altmetric client the DOI is the primary identifier, and is the method used to retrieve Altmetric data from the API. Very nearly all joins between datasets are achieved using the doi field (the exception being the join between mentions and authors).

### type

**Data type: character**

This is a category determined by Altmetric themselves, and appears to always be 'article'.

```{r Grouping by type, echo=TRUE}

  summarise(
    group_by(
      articles, 
      type), 
    total = n()
    )

```

